Lossy compression of DNA sequencing quality data

### 2016-04-12 bmc_manuscript/manuscipt.tex
Created a merged version of the two manuscripts that Chris and I independently created.

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
   noteref={n1},                        % id's of article notes, if any
   email={jane.e.doe@cambridge.co.uk}   % email address
]{\inits{CMH}\fnm{Christopher M} \snm{Hill}}
\author[
   addressref={aff3,aff5},
   email={szolek@informatik.uni-tuebingen.de}
]{\inits{AS}\fnm{Andr\'{a}s} \snm{Szolek}}
\author[
   addressref={aff4,aff5},
   email={mohamed.el-hadidi@uni-tuebingen.de}
]{\inits{MEH}\fnm{Mohamed} \snm{El Hadidi}}
\author[
   addressref={aff2},
   corref={aff2},                       % id of corresponding address, if any
   email={mike@umiac.umd.edu}
]{\inits{MPC}\fnm{Michael P} \snm{Cummings}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{Department of Computer Science, University of Maryland}, % university, etc
  \city{College Park, Maryland},                              % city
  \postcode{20742}                                % post or zip code
  \cny{USA}                                    % country
}
\address[id=aff2]{%                           % unique id
  \orgname{Center for Bioinformatics and Computational Biology, University of Maryland}, % university, etc
  \city{College Park, Maryland},                              % city
  \postcode{20742}                                % post or zip code
  \cny{USA}                                    % country
}
\address[id=aff3]{%
  \orgname{Department of Applied Bioinformatics, Center for
  Bioinformatics, Quantitative Biology Center, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}
\address[id=aff4]{%
  \orgname{Department of Algorithms in Bioinformatics, Center for
  Bioinformatics, Quantitative Biology Center, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}
\address[id=aff5]{%
  \orgname{Department of Computer Science, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}

@@

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Current address: \ldots} % note, connected to author
\end{artnotes}

@@

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

@@

Previous literature on sequence data compression has largely focused
on lossless compression of base calls~\cite{Deorowicz:2013hq,
  Giancarlo:2009fk,Giancarlo:2014rw, Nalbantoglu:2010uq,Zhu:2013qr},
although some recent work has addressed compression of quality
values~\cite{Canovas:2014fr,Hach:2012ys,
  janin2013adaptive,Kozanitis:2011kl,Ochoa:2013rt,Tembe:2010ys,
  Wan:2012kq,DBLP:conf/recomb/YuYB14,zhou2014compression}. Among the
several challenges for compression of read data is dealing with
different error profiles resulting from differences in underlying
chemistries, signal detection and processing mechanisms, inherent
biases, and other idiosyncratic properties of distinct high-throughput
sequencing technologies.

@@

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{compression}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

>>>>

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
%   noteref={n1},                        % id's of article notes, if any
   email={chrismh@uw.edu}   % email address
]{\inits{CMH}\fnm{Christopher M} \snm{Hill}}
\author[
   addressref={aff4,aff6},
   email={szolek@informatik.uni-tuebingen.de}
]{\inits{AS}\fnm{Andr\'{a}s} \snm{Szolek}}
\author[
   addressref={aff5,aff6},
   email={mohamed.el-hadidi@uni-tuebingen.de}
]{\inits{MEH}\fnm{Mohamed} \snm{El Hadidi}}
\author[
   addressref={aff3},
   corref={aff2},                       % id of corresponding address, if any
   email={mike@umiac.umd.edu}
]{\inits{MPC}\fnm{Michael P} \snm{Cummings}}

@@

\address[id=aff1]{%                           % unique id
  \orgname{Department of Computer Science, University of Maryland}, % university, etc
  \city{College Park, Maryland},                              % city
  \postcode{20742}                                % post or zip code
  \cny{USA}                                    % country
}
\address[id=aff2]{%                           % unique id
  \orgname{Department of Genome Science, University of Washington School of Medicine},                    %
  \city{Seattle, Washington},                              % city
  \postcode{98195},                                % post or zip code
  \cny{USA}                                    % country
}
\address[id=aff3]{%                           % unique id
  \orgname{Center for Bioinformatics and Computational Biology, University of Maryland}, % university, etc
  \city{College Park, Maryland},                              % city
  \postcode{20742}                                % post or zip code
  \cny{USA}                                    % country
}
\address[id=aff4]{%
  \orgname{Department of Applied Bioinformatics, Center for
  Bioinformatics, Quantitative Biology Center, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}
\address[id=aff5]{%
  \orgname{Department of Algorithms in Bioinformatics, Center for
  Bioinformatics, Quantitative Biology Center, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}
\address[id=aff6]{%
  \orgname{Department of Computer Science, University of T\"{u}bingen},
  \street{Sand 14},
  \postcode{72076}
  \city{T\"{u}bingen},
  \cny{Germany}
}

@@

%\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Current address: \ldots} % note, connected to author
%\end{artnotes}

@@

\begin{abstract} % abstract
\parttitle{Motivation:} %if any
As the cost of sequencing continues to decrease, the rate of sequence
data production is increasing, placing greater demands on storing and
transferring these vast amount of data. Most methods of sequencing
data compression focus on compressing nucleotide information without
any loss of information. Quality data, however, have different
properties than nucleotide data, and methods compressing nucleotide
sequences efficiently do not perform as well on quality
values. Furthermore, although lossless representation might be
necessary for nucleotide sequences, it is not an essential requirement
for quality values.

Previous studies of quality value compression have mostly focus on
minimizing the loss of information with less emphasis on the effects
on bioinformatic analyses. In this paper, we evaluate several
different compression methods for quality values, and assess the
resulting impacts on common bioinformatic analyses using sequence read
data: quality control, genome assembly, and alignment of short reads
to a reference sequence.

\parttitle{Results:} %if any
Lossy compression of quality information can greatly decrease memory
requirements, and our results demonstrate
that some compression methods can result in transformed quality values
that are quite useful, and in some cases advantageous, compared to
original uncompressed values.
\end{abstract}

@@

Previous literature on sequence data compression has largely focused
on lossless compression of base calls~\cite{Deorowicz:2013hq,
  Giancarlo:2009fk,Giancarlo:2014rw, Nalbantoglu:2010uq,Zhu:2013qr},
although some recent work has addressed compression of quality
values~\cite{Canovas:2014fr,Hach:2012ys,
  janin2013adaptive,Kozanitis:2011kl,Ochoa:2013rt,Tembe:2010ys,
  Wan:2012kq,DBLP:conf/recomb/YuYB14,zhou2014compression,Malysa01102015}. Among the
several challenges for compression of read data is dealing with
different error profiles resulting from differences in underlying
chemistries, signal detection and processing mechanisms, inherent
biases, and other idiosyncratic properties of distinct high-throughput
sequencing technologies.

@@

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
%\bibliographystyle{vancouver} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{compression}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}
